BeeModeling

### 2022-07-26 Fowler-GLOBI Modeling/GLOBI-Fowler Data Prep-reformatted.Rmd
starting script

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/colleen/Dropbox/Fall_2014/ucsb/Bee-Specialization-Modeling/")
knitr::opts_knit$set(error=T)

@@

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('modeling_data/interactions_dup.csv')

@@

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}
```

```{r}
interactions_b[1:5,]
```

@@

table_test$specialist <- table_test$west_specialist + table_test$east_specialist + table_test$cent_specialist
table_test$specialist <- as.integer(table_test$specialist > 0)
table_test$specialist[is.na(table_test$specialist)] <- 0

write_csv(table_test, 'modeling_data/fowler')
```

```{r}
plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)
plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/colleen/Dropbox/Fall_2014/ucsb/Bee-Specialization-Modeling/")
knitr::opts_knit$set(error=T)
select = dplyr::select

@@

```{r, echo = False}
options(max.print=1000000)
interactions_b <- read.csv('modeling_data/interactions_dup.csv')
ms <- read_xlsx('modeling_data/DATA_pollen-diet-breadth_list_12-5-21_ALR.xlsx', sheet = 3) %>%
  rename_all(tolower) %>%
  mutate(scientificName = paste0(genus, ' ',sub('.*\\. ','',species))) %>%
  select(scientificName,everything()) %>%
  rename(diet_breadth=`redefined pollen diet breadth`)
fowler <- read_csv("modeling_data/fowler_hostplants.csv")

#who are the specialsits?
#let's say specialist the MS list and everything on fowler list that's missing from ms list
ms_oligo = ms$scientificName[ms$diet_breadth %in% c('Oligolectic','Monolectic')]
add_from_fowler = unique(fowler$scientificName[!fowler$scientificName %in% ms$scientificName])
specialists = c(ms_oligo,add_from_fowler)

@@

# for (c in colnames(interactions_b)) {
#   interactions_b[[c]] <-  tolower(interactions_b[[c]])
# }
```

```{r}
#how many species in the missouri state data are in the globi data?
head(interactions_b)
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms$scientificName)]) #800 species in the ms data

#how many are specialists
ms_specialists = ms %>% filter(diet_breadth != 'Polylectic')
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_specialists$scientificName)]) #503 specialist species

#how many are generalists
ms_generalists = ms %>% filter(diet_breadth == 'Polylectic')
n_distinct(interactions_b$sourceTaxonName[interactions_b$sourceTaxonName  %in% tolower(ms_generalists$scientificName)]) #296 generalist species

```{r}
interactions_b[1:5,1:5]
```

Let's do some reformatting
```{r}
capitalize = function(a_string){

  beginning = toupper(substr(a_string,1,1))
  ending = substr(a_string,2,nchar(a_string))

  return(paste0(beginning,ending))

}

globi_summ = interactions_b %>% 
  mutate(scientificName = capitalize(sourceTaxonSpeciesName), 
         plant_family = capitalize(targetTaxonFamilyName), 
         plant_genus = capitalize(targetTaxonGenusName)) %>%
  filter(scientificName != '' & plant_family !='' & plant_genus != "") %>% 
  group_by(scientificName, plant_genus, plant_family,) %>% 
  summarize(n_genus = n()) %>% ungroup()

how many specialist bees do we have if we filter out interactions with <5 observations?
```{r}
globi_5n = globi_summ %>% filter(n_genus >= 5)
n_distinct(globi_5n$scientificName[globi_5n$scientificName %in% ms_specialists$scientificName]) #only 360

#how many other species
n_distinct(globi_5n$scientificName[!globi_5n$scientificName %in% ms_specialists$scientificName]) # 730, so about 1000 total

```
Calculate number of interaction partners

```{r}
#calculate the number of plant families each bee species is observed visiting
total_obs = globi_summ %>% group_by(scientificName) %>% summarize(n=sum(n_genus))

# degree_by_fam 
degree_by_fam = globi_summ %>% group_by(scientificName,plant_family) %>%
  summarize(n_family = sum(n_genus)) %>%
  left_join(total_obs) %>%
  mutate(freq_squared = (n_family/n)^2) %>% group_by(scientificName,n) %>%
  summarize(degree_family = n_distinct(plant_family), 
            degree5_family = sum(n_family >= 5), weighted_degree_fam = sum(freq_squared))# %>%

#calculate the number of plant genera each bee species is observed visiting
degree_by_genus = globi_summ %>%
  left_join(total_obs) %>%
  mutate(freq_squared = (n_genus/n)^2) %>% 
  group_by(scientificName) %>%
  summarize(degree_genus = n_distinct(plant_genus), 
            degree5_genus = sum(n_genus >= 5), weighted_degree_genus = sum(freq_squared))

globi_degree = degree_by_fam %>% left_join(degree_by_genus) %>%
  mutate(diet_breadth = ifelse(scientificName %in% specialists,'specialist','generalist'))

write_csv(globi_degree,'modeling_data/globi_degree.csv')

@@

table_test$specialist <- table_test$west_specialist + table_test$east_specialist + table_test$cent_specialist
table_test$specialist <- as.integer(table_test$specialist > 0)
table_test$specialist[is.na(table_test$specialist)] <- 0
getwd()

# write_csv(table_test, 'modeling_data/fowler')
```
```{r reformat the table to include host plant data}
fowler_table = eastern_table %>% mutate(region = 'east') %>%
  bind_rows(central_table %>% mutate(region='central')) %>%
  bind_rows(western_table %>% mutate(region = 'west')) %>%
  rename(host_plant = Host.plant.Family..Tribe..Genera) %>%
  rename_all(tolower) %>%
  mutate(scientificName = gsub(" *\\(.*?\\) *", "", genus..subgenus..species)) %>%
  select(scientificName,authority,status,host_plant,region) %>%
  #remove astericks from some of the scientific names:
  mutate(scientificName = ifelse(grepl("\\*",scientificName),sub("\\*.*","",scientificName),scientificName))

# write_csv(fowler_table,'modeling_data/fowler_hostplants.csv')
```

```{r}
plotdf <- merge(beeCite, table_test, by = 'interestBee', all.x = TRUE)

nrow(plotdf)

plotdf$specialist <- plotdf$west_specialist + plotdf$east_specialist + plotdf$cent_specialist
plotdf$specialist <- as.integer(plotdf$specialist > 0)
plotdf$specialist[is.na(plotdf$specialist)] <- 0
